---
title: "Individual Research Meeting"
subtitle: "개인 프로젝트 미팅" 
Week: 15
format: html
slide-format: revealjs
---

[Weekly content](https://changjunlee.com/teaching/grad_stat/weekly_2/)

<br>

[개별 연구 미팅 신청 탭](https://docs.google.com/spreadsheets/d/1fc825-IgQMe1gbs5sEKhtBOPbYBN3gVCPjgXvXOwXj0/edit?usp=sharing)

<bR>

## Structural Equation Modeling: Foundations and Applications

<br>

### **Introduction**

Structural Equation Modeling (SEM) is a comprehensive statistical methodology that combines factor analysis and multiple regression techniques to examine complex relationships among observed and latent variables. Over the past few decades, SEM has become a powerful tool for researchers across disciplines such as social sciences, psychology, education, business, and health sciences. This chapter introduces the theoretical underpinnings of SEM, its key components, and practical applications, with an emphasis on latent variables and path analysis.

<bR>

### **Key Concepts in SEM**

#### **1. Observed and Latent Variables**

-   **Observed Variables**: These are directly measured variables, often referred to as indicators. For example, responses to survey items like "I enjoy learning" and "I find studying rewarding" can serve as observed variables for the latent construct "Motivation."

-   **Latent Variables**: These are unobserved constructs that are inferred from observed variables. SEM excels in estimating latent variables while accounting for measurement error.

#### **2. Measurement Error**

Measurement error is the discrepancy between the observed value and the true value of a variable. SEM explicitly incorporates measurement error into its models, which sets it apart from traditional regression approaches.

#### **3. Path Diagrams**

Path diagrams visually represent SEM models. Key components include:

-   **Circles**: Represent latent variables.

-   **Squares**: Represent observed variables.

-   **Arrows**: Depict causal relationships or associations.

<br>

### **Components of SEM**

#### **1. Measurement Model**

The measurement model specifies the relationships between latent variables and their observed indicators. It answers the question: *How well do the observed variables measure the underlying construct?*

**Example**: Motivation as a latent variable might be measured by three observed items:

$$
Motivation \sim Item_1 + Item_2 + Item_3
$$

​<br>

#### **2. Structural Model**

The structural model specifies the relationships between latent variables (or between latent and observed variables). It answers the question: *What are the causal relationships between constructs?*

**Example**: Testing whether motivation mediates the relationship between hours studied and test performance:

$$
Test\_Score = Motivation + Hours\_Studied
$$

$$
Motivation = Hours\_Studied
$$

<br>

### **Steps in Conducting SEM**

#### **Step 1: Define the Model**

Develop a theoretical model based on prior research or hypotheses. Clearly specify the relationships among variables in a path diagram.

#### **Step 2: Specify the Model**

Translate the path diagram into a set of equations. This involves defining the measurement and structural models.

#### **Step 3: Collect Data**

Ensure the dataset includes sufficient sample size and all observed indicators for latent variables.

#### **Step 4: Estimate the Model**

Use software such as R (`lavaan` package), AMOS, Mplus, or LISREL to estimate the parameters.

#### **Step 5: Evaluate Model Fit**

Examine fit indices and modify the model, if necessary, based on theoretical justifications.

#### **Step 6: Interpret the Results**

Interpret path coefficients, factor loadings, and fit indices in the context of your research question.

<Br>

## **Cronbach's Alpha: A Measure of Reliability**

#### **Introduction**

Cronbach's Alpha (α) is a widely used statistic in research to assess the **internal consistency reliability** of a set of items or indicators. It is particularly useful for measuring the reliability of scales, tests, or questionnaires in social sciences, psychology, education, and other fields. Reliability refers to the extent to which a scale produces consistent results across repeated measurements or multiple items measuring the same construct.

### **Conceptual Foundation**

#### **Internal Consistency**

Internal consistency evaluates how well the items in a scale measure the same underlying construct. If items are highly correlated, it suggests they are measuring the same concept, which contributes to higher reliability.

For example:

-   A scale measuring **motivation** might include items like:

    -   "I enjoy learning new things."

    -   "I am motivated to achieve my goals."

    -   "I find studying rewarding."

If these items are highly interrelated, the scale has good internal consistency.

#### **Definition of Cronbach's Alpha**

Cronbach's Alpha quantifies internal consistency as a value between 0 and 1. Higher values indicate better reliability. The formula for α is:

$$
\alpha = \frac{k}{k - 1} \left(1 - \frac{\sum s_i^2}{s_t^2}\right)
$$

Where:

-   k: Number of items in the scale.

-   $s_i^2$​: Variance of each individual item.

-   $s_t^2$​: Variance of the total scale score.

<br>

### **Interpreting Cronbach's Alpha**

-   α \> 0.90: Excellent reliability (possibly redundant items).

-   0.80 ≤ α \< 0.90: Good reliability.

-   0.70 ≤ α \< 0.80: Acceptable reliability.

-   0.60 ≤ α \< 0.70: Questionable reliability.

-   α \< 0.60: Poor reliability (likely indicates issues with item quality or scale design).

<Br>

### **Assumptions of Cronbach's Alpha**

1.  **Unidimensionality**: The scale should measure a single construct.

2.  **Equal Variance Contribution**: Each item should contribute equally to the total score (violations can lead to under- or overestimation of reliability).

3.  **Correlation Between Items**: Items should be positively correlated.

<Br>

### **Practical Example: Calculating Cronbach's Alpha in R**

#### **Simulating a Dataset**

```{r}
# Simulate data for a 5-item motivation scale
set.seed(123)
n <- 100  # Number of respondents
motivation_data <- data.frame(
  Q1 = rnorm(n, mean = 4, sd = 0.8),  # Item 1
  Q2 = rnorm(n, mean = 4.2, sd = 0.7),  # Item 2
  Q3 = rnorm(n, mean = 3.9, sd = 0.9),  # Item 3
  Q4 = rnorm(n, mean = 4.1, sd = 0.6),  # Item 4
  Q5 = rnorm(n, mean = 4.0, sd = 0.8)   # Item 5
)
motivation_data

```

```{r}
# Install and load the psych package
# install.packages("psych")
library(psych)

# Calculate Cronbach's Alpha
alpha_result <- psych::alpha(motivation_data)
print(alpha_result)

```

#### **Output Interpretation**

The output includes:

-   **Raw Alpha**: The Cronbach's Alpha for the current scale.

-   **Standardized Alpha**: Adjusted for standardized item variances.

-   **Item-Total Correlations**: Correlation of each item with the total scale score.

-   **Alpha If Deleted**: The α value if a specific item is removed.

<br>

### **Improving Reliability**

1.  **Increase the Number of Items**: Add more items that measure the same construct.

2.  **Refine Items**: Ensure items are clear, relevant, and unambiguous.

3.  **Remove Low-Quality Items**: Exclude items that weaken the overall reliability (α).

<br>

## **Confirmatory Factor Analysis (CFA)**

### **ntroduction**

Confirmatory Factor Analysis (CFA) is a statistical technique used to test whether a hypothesized measurement model fits the observed data. Unlike Exploratory Factor Analysis (EFA), which identifies underlying factor structures without prior assumptions, CFA is **hypothesis-driven** and requires the researcher to specify the relationships between observed variables and latent factors before analysis.

CFA is widely used in psychology, education, and social sciences to validate measurement scales, confirm theoretical constructs, and ensure the reliability of instruments.

<br>

### **Key Concepts in CFA**

#### **1. Latent Variables**

Latent variables are unobserved constructs that are inferred from multiple observed variables. For example:

-   **Motivation** might be a latent variable inferred from survey items like "I enjoy studying" and "I set academic goals."

#### **2. Observed Variables (Indicators)**

These are measurable variables that serve as proxies for the latent variable. In CFA, each observed variable is expected to load on a specific latent factor.

#### **3. Factor Loadings**

Factor loadings quantify the relationship between observed variables and their underlying latent factor. Higher loadings indicate stronger relationships.

#### **4. Measurement Errors**

CFA explicitly models measurement error for each observed variable, improving the accuracy of parameter estimates compared to traditional methods.

<br>

### **The CFA Model**

The CFA model is typically represented as:

$$
X = \Lambda \xi + \delta
$$

Where:

-   X: Vector of observed variables.

-   Λ: Matrix of factor loadings.

-   ξ: Vector of latent variables.

-   δ: Vector of measurement errors.

<bR>

### **Steps in Conducting CFA**

#### **Step 1: Specify the Model**

Develop a hypothesized factor structure based on theory or prior research. This includes:

-   Defining the number of factors.

-   Assigning observed variables to latent factors.

-   Specifying relationships between factors (if applicable).

#### **Step 2: Collect Data**

Ensure that the sample size is sufficient. A common rule of thumb is **10-20 observations per variable**.

#### **Step 3: Estimate the Model**

Fit the model using software such as R (`lavaan` package), AMOS, Mplus, or LISREL. Maximum Likelihood (ML) is the most commonly used estimation method.

#### **Step 4: Assess Model Fit**

Evaluate the fit of the hypothesized model using various fit indices (see below).

#### **Step 5: Refine the Model**

Modify the model if necessary, based on theoretical justifications and statistical outputs.

<br>

### **Model Fit Indices in CFA**

Several indices are used to assess the goodness of fit of a CFA model:

1.  **Chi-Square**:

    -   Tests the null hypothesis that the model fits the data perfectly.

    -   Smaller, non-significant values indicate a good fit.

    -   Sensitive to sample size.

2.  **CFI (Comparative Fit Index)**:

    -   Compares the fit of the hypothesized model to a null model.

    -   Values \> 0.90 indicate good fit; \> 0.95 indicates excellent fit.

3.  **TLI (Tucker-Lewis Index)**:

    -   Adjusts for model complexity.

    -   Values \> 0.90 indicate good fit.

4.  **RMSEA (Root Mean Square Error of Approximation)**:

    -   Measures the discrepancy between the model and the data per degree of freedom.

    -   Values \< 0.08 indicate acceptable fit; \< 0.05 indicates excellent fit.

5.  **SRMR (Standardized Root Mean Square Residual)**:

    -   Measures the average discrepancy between observed and predicted correlations.

    -   Values \< 0.08 indicate a good fit.

<br>

### **Example of CFA in R**

Here, we validate a hypothetical 3-factor model with 9 observed variables.

#### **Simulating the Data**

```{r}
# Simulate data for a 3-factor model
set.seed(123)
n <- 300

# Latent variables
Factor1 <- rnorm(n, mean = 5, sd = 1)
Factor2 <- rnorm(n, mean = 3, sd = 1)
Factor3 <- rnorm(n, mean = 4, sd = 1)

# Observed variables
Item1 <- 0.8 * Factor1 + rnorm(n, sd = 0.5)
Item2 <- 0.7 * Factor1 + rnorm(n, sd = 0.5)
Item3 <- 0.9 * Factor1 + rnorm(n, sd = 0.5)

Item4 <- 0.8 * Factor2 + rnorm(n, sd = 0.5)
Item5 <- 0.7 * Factor2 + rnorm(n, sd = 0.5)
Item6 <- 0.9 * Factor2 + rnorm(n, sd = 0.5)

Item7 <- 0.8 * Factor3 + rnorm(n, sd = 0.5)
Item8 <- 0.7 * Factor3 + rnorm(n, sd = 0.5)
Item9 <- 0.9 * Factor3 + rnorm(n, sd = 0.5)

data <- data.frame(Item1, Item2, Item3, Item4, Item5, Item6, Item7, Item8, Item9)
data
```

Defining and Running the CFA Model

```{r}
# Load lavaan package
library(lavaan)

# Define the CFA model
model <- '
  Factor1 =~ Item1 + Item2 + Item3
  Factor2 =~ Item4 + Item5 + Item6
  Factor3 =~ Item7 + Item8 + Item9
'

# Fit the model
fit <- cfa(model, data = data)

# Summary with standardized estimates
summary(fit, fit.measures = TRUE, standardized = TRUE)

```

#### **Output**

The output provides:

-   **Factor Loadings**: Should generally be \> 0.5.

-   **Fit Indices**: Evaluate overall model fit.

-   **Residuals**: Indicate discrepancies between observed and predicted values.

<Br>

### **Applications of CFA**

1.  **Scale Validation**: Confirm whether a set of items measures the intended construct.

2.  **Construct Validation**: Test theoretical constructs (e.g., motivation, anxiety).

3.  **Measurement Invariance**: Assess whether a scale functions equivalently across groups (e.g., gender, culture).

<br>
